---
title: "congress-data"
author: Mark Barrett
editor: source
format:
  html: 
    toc: true
    df-print: paged
---

# Getting Set Up

Start R project named `congress-data`

-   Select git repo

-   Select `renv` because I like to complicate things (but the reproducibility!)

Try to send the repo to GitHub with `git push`. But I haven't added the repo on GitHub

Use `git remote add origin [YourGitURL]`. Commit initial files, then try pushing again. The terminal errors and gives a `--set-upstream` command that solved the issue?

## Libraries

```{r}
#| label: load-libs
#| message: false
#| warning: false
library(tidyverse)
library(here)
library(lubridate)
library(skimr)
```

# Legislative data

Source: govInfo

Record of legislation introduced in the House and Senate of the US Congress. In this case, I am looking at the 117th Congress which was in power from January 2021 - January 2023.

```{r}
# Load latest version of clean data
clean_data = fs::dir_info(here::here("data", "cleaned"))
latest_file = clean_data[which.max(clean_data$change_time), "path", drop = T]

all_bills = readRDS(latest_file)

```

```{r}
skim(all_bills)
```

How much legislation goes through congress? In our dataset, we have `r nrow(all_bills)` pieces of legislation. We can break this out by the number of bills originating in the House compared to the Senate.

```{r}
all_bills |> 
  count(origin_chamber)
```

Most legislation at least begins in the House.

How much legislation is considered on a given day?

```{r}
# Unnest actions
# Count distinct bills by action date
```

How much legislation becomes law?

```{r}
all_bills = all_bills |> 
  mutate(bill_passed = map_lgl(actions, \(bill_actions) "BecameLaw" %in% bill_actions$type))

bills_passed = all_bills |> 
  filter(bill_passed)

nrow(bills_passed)
```

Check the bill type though:

```{r}
bills_passed |> 
  count(type, sort = T)
```

The 117th Congress was unusual in that the Democrats had control of all three branches of government.
Of the 365 laws that were passed, how many were sponsored by Republicans?

```{r}
bills_passed |> 
  unnest(sponsors) |> 
  count(party)
```


## Policy Areas

What kind of topics are the laws related to? What about all legislation proposed?

```{r}
bills_passed |> 
  unnest(policy_areas, keep_empty = T) |>
  # Count bills by policy areas
  group_by(policy_areas) |>
  summarise(bills = n_distinct(bill_id)) |>
  arrange(desc(bills))
```

```{r}
all_bills |> 
  unnest(policy_areas, keep_empty = T) |>
  # Count bills by policy areas
  group_by(policy_areas) |>
  summarise(n = n(),
            bills = n_distinct(bill_id),
            bills_passed = sum(bill_passed)) |>
  arrange(desc(bills))
```


Let's take a look at laws related to environmental protection.

```{r}
bills_passed_ep = bills_passed |> 
  unnest(policy_areas) |> 
  filter(policy_areas == "Environmental Protection")

View(bills_passed_ep)
```

## Committees

```{r}
committees = all_bills |> 
  unnest(committees, names_sep = "_", keep_empty = T) |> 
  unnest(committees_activities, names_sep = "_", keep_empty = T)
  
```

How active are different committees?

```{r}
committees |> 
  group_by(committees_chamber, committees_name, bill_passed) |> 
  summarise(bills = n_distinct(bill_id), .groups = "drop")
```

```{r}
committees |> 
  count(committees_chamber, committees_type, committees_name, committees_activities_name) |> 
  pivot_wider(names_from = committees_activities_name, values_from = n)
```


-   How many different bills do they consider?

-   How many hearings do they hold?

-   How many bills do they report?
