---
title: "Using `congress-data`"
author: Mark Barrett
editor: source
format:
  html: 
    toc: true
    df-print: paged
---

# Libraries

```{r}
#| label: load-libs
#| message: false
#| warning: false
library(tidyverse)
library(here)
library(lubridate)
library(skimr)
```

# Legislative data

Source: govInfo's [Bill Status bulk repository](https://www.govinfo.gov/bulkdata/BILLSTATUS)

I collected this data for the 117th Congress (2021-2023) by downloading the XML files and parsing them into a tabular format (Link to data collection article coming soon!) using R. Below I load this data file:

```{r}
#| label: load-data
#| cache: true
# Load latest version of clean data
clean_data = fs::dir_info(here::here("data", "cleaned"))
latest_file = clean_data[which.max(clean_data$change_time), "path", drop = T]

all_bills = readRDS(latest_file)

```

## What even is legislation?

Before we explore the legislative data, it's important to understand how the legislative process actually works. Congress has some videos that give an [overview of the legislative process](https://www.congress.gov/legislative-process) in the style of a mandatory corporate compliance training. [How Our Laws Are Made](https://www.congress.gov/help/learn-about-the-legislative-process/how-our-laws-are-made) is a more thorough outline of the federal law-making process, written by former parliamentarian John Sullivan.

So what happens between the spark of an idea and the presidential signature that enacts it into law? Simply put, the idea is written as a bill to be introduced to Congress and referred to committees of legislators that focus on the policy area(s) of the bill.\
The committee(s) can report the bill for *consideration*[^1] or the whole chamber can decide to vote on it immediately. If the bill is passed by one chamber of Congress, it moves on to the opposite chamber which employs a similar process and finally the president signs the bill into law.

[^1]: Committees typically report a bill for consideration to move the bill to recommend the chamber votes on the bill. They can also report a bill unfavourably but as we'll see that doesn't happen too often.

How is this process recorded in data and can we learn anything about the institutional decision-making by exploring the data?

::: {.callout-note collapse="true"}
## Get a glimpse of the data

`glimpse()` shows us each variable, their type, and a few example data points.

```{r}
glimpse(all_bills)
```
:::

```{r}
#| label: skim-df
#| eval: false
#| include: false
#Get a summary of the data first.
skim(all_bills)
```

In addition to administrative information like the bill number, title, and and sponsors, Bill Status contains the committees that a bill was referred to, actions taken by different congressional bodies, and the policy area classifications of the bill.

To start, how many pieces of legislation are presented to Congress by our representatives?\
In the 117th congress, there were **`r scales::comma(nrow(all_bills))`** bills and resolutions introduced throughout its two legislative sessions. 
How much legislation is introduced in either chamber?

```{r}
#| label: chamber-bills
all_bills |> 
  # Count the number of bills by their chamber of origin
  count(origin_chamber, wt = n_distinct(bill_id))
```

Nearly double the amount of legislation was introduced in the House as in the Senate. 
The House does have over four times the number of legislators as the Senate and also has the unique ability to introduce bills which raise revenue (e.g. taxes).
With all these bills being introduced, how many become law?

```{r}
#| label: bills-passed
all_bills = all_bills |> 
  # Search each bill's actions for a "BecameLaw" action type
  mutate(bill_passed = map_lgl(actions, \(bill_actions) "BecameLaw" %in% bill_actions$type))

bills_passed = all_bills |> 
  # Keep only passing bills
  filter(bill_passed)

nrow(bills_passed)
```

Only `r nrow(bills_passed)`.

## Sponsors

The 117th Congress was unusual in that the Democrats had control of all three branches of government. Of the 365 laws that were passed, how many were sponsored by Republicans?

```{r}
#| label: sponsor-party-check
bills_passed |> 
  select(bill_id, sponsors) |> 
  # Unnest sponsors
  unnest(sponsors, keep_empty = T) |> 
  # Count the party of bills' sponsors
  count(party) |> 
  mutate(pct = scales::percent(n/sum(n)))
```

How much legislation was introduced by Democrats and Republicans in each chamber?

```{r}
all_bills |> 
  select(bill_id, origin_chamber, sponsors) |> 
  # Unnest sponsors data
  unnest(sponsors, keep_empty = T) |> 
  # Count unique bills by chamber and party
  group_by(origin_chamber, party) |> 
  summarise(bills = n_distinct(bill_id),
            .groups = "drop") |> 
  # Pivot chamber to the columns
  pivot_wider(names_from = origin_chamber, values_from = bills, values_fill = 0) |> #, values_fill = "0", values_fn = scales::comma)
  # Calculate percentages
  mutate(`House %` = House/sum(House),
         `Senate %` = Senate/sum(Senate),
         # Format columns
         across(c(House, Senate), scales::comma),
         across(contains("%"), scales::percent, accuracy = .1)) |> 
  dplyr::relocate(`House %`, .after = House)
```


So roughly a third of the laws passed were introduced by Republicans. Lawmakers can also *co-sponsor* bills indicating their support for it. How often do bills have co-sponsors? And how often do they have co-sponsors from the opposite party?

```{r}
all_bills |> 
  select(bill_id, cosponsors) |> 
  unnest(cosponsors, keep_empty = T) |> 
  # Count sponsors per bill
  group_by(bill_id) |> 
  summarise(cosponsors = sum(!is.na(bioguide_id))) |> 
  summarise(across(cosponsors, list(min = min, median = median, mean = mean, max = max), na.rm=T))
```

```{r}
# Find sponsors and their parties
all_bills |> 
  select(bill_id, sponsors, cosponsors) |> 
  unnest(sponsors, keep_empty = T, names_sep = "_") |> 
  select(bill_id, sponsors_full_name, sponsors_party, cosponsors) |> 
  # Find cosponsors and their parties
  unnest(cosponsors, keep_empty = T, names_sep = "_") |> 
  select(bill_id, sponsors_full_name, sponsors_party, cosponsors_full_name, cosponsors_party) |> 
  group_by(sponsors_party, cosponsors_party) |> 
  summarise(bills = n_distinct(bill_id))
  
```

How many of the bills that became law were introduced by either party?

```{r}
bills_passed |> 
  # Unnest sponsors data
  select(bill_id, origin_chamber, sponsors) |> 
  unnest(sponsors, keep_empty = T) |> 
  group_by(origin_chamber, party) |> 
  summarise(bills = n_distinct(bill_id)) |> 
  pivot_wider(names_from = origin_chamber, values_from = bills)
```


That is a very small number compared to how many bills are introduced. Now it's important to note that only certain kinds of bills *can* become law - (regular old) bills and joint resolutions.

```{r}
filter(all_bills, type %in% c("HR", "HJRES", "S", "SJRES")) |> 
  nrow()
```

So around 2-3% become law.

```{r}
#| eval: false
#| include: false
# Check the bill type though, are any simple/concurrent resolutions in there?
bills_passed |> 
  count(type, sort = T)
```


Questions

-   Do successful bills get sponsored by a small number of legislators?

-   What kind of bills did Democrats sponsor compared to Republicans?


## Policy Areas

What policy areas are these new laws related to? What about all legislation proposed?

```{r}
#| label: policy-bill-count
all_bills |> 
  unnest(policy_areas, keep_empty = T) |> # nrow() # check row count
  # Count bills by policy areas
  group_by(policy_areas) |>
  summarise(bills = n_distinct(bill_id),
            bills_passed = sum(bill_passed)) |>
  arrange(desc(bills_passed))
```

## Committees

How many bills are referred to different committees?

```{r}
#| label: committee-count
#| column: page
committees = all_bills |> 
  unnest(committees, names_sep = "_", keep_empty = T) |> 
  unnest(committees_activities, names_sep = "_", keep_empty = T)

committees |> 
  group_by(committees_chamber, committees_name) |> 
  summarise(bills = n_distinct(bill_id), .groups = "drop")
  
```

```{r}
#| eval: false
#| include: false
# What's happening with rows that are missing committees chamber and name?
committees |> 
  filter(is.na(committees_chamber), is.na(committees_name))
```

What activities do committees engage in?

```{r}
#| label: committee-activity-count
#| column: page
committees |> 
  # Count committee activities on bills
  count(committees_chamber, committees_name, committees_activities_name, wt = n_distinct(bill_id)) |> 
  pivot_wider(names_from = committees_activities_name, values_from = n, values_fill = 0) |> 
  # Sort activities by their (rough) order of occurrence
  select(committees_name, committees_chamber, 
         `Referred to`, `Hearings by`, `Markup by`, `Reported by`, `Reported original measure`, `Discharged from`, `NA`)
```

```{r}
# Why does this bill have an action occur after it became law?
bills_passed |> 
  select(bill_id, contains("action")) |> 
  filter(bill_id == "HR-1276") |> 
  unnest(actions) |> 
  arrange(action_date)
```

# Questions

-   How many different bills do committees consider?

-   How many hearings do committees hold?

-   How many bills do committees report?

-   What actions occur on bills that become law? How does this differ from bills that do not?

-   How many bills that become law received a committee report? How many had hearings?

- How long does it take for bills to become law?

- Where happens to bills that don't become law?



::: {.callout-note collapse="true" appearance="minimal"}
## Project setup

Start R project named `congress-data`

-   Select git repo

-   Select `renv` because I like to complicate things (but the reproducibility!)

Try to send the repo to GitHub with `git push`. But I haven't added the repo on GitHub

Use `git remote add origin [YourGitURL]`. Commit initial files, then try pushing again. The terminal returns an error and provides a `--set-upstream` command that solves the issue?
:::